Cultural Transmission Can Explain the Evolution of Cooperation

### 2020-08-03 ms/main.tex
added coexistence figure; revised figure captions and formating

We define the initial frequency as $\tilde{p}_0$ and the cost boundaries
\begin{equation}\begin{aligned}
\gamma_1 = \frac{b \alpha T_A + (T_A - T_B)}{1-T_B}, \quad
\gamma_2 = \frac{b \alpha T_B + (1+b) (T_A - T_B)}{1-T_B}.
\end{aligned}\end{equation}
Applying eqs.~\ref{eq:vert_hori_equilibrium}, \ref{eq:vert_hori_global_condition1}, and \ref{eq:vert_hori_global_condition2} we can summarize the possible outcomes:
%\begin{enumerate} % p* version
%\item Fixation of cooperation,
%	if $T_A>T_B$ and $0<\tilde{p}^*$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0>\tilde{p}^*$, or
%	if $T_A<T_B$ and $1<\tilde{p}^*$.
%\item Global fixation of defection,
%	if $T_A>T_B$ and $\tilde{p}^*>1$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0<\tilde{p}^*$, or
%	if $T_A<T_B$ and $\tilde{p}^*<0$
%\item Co-existence of both phenotypes at $\tilde{p}^*$, 
%	if $T_A<T_B$ and $0<\tilde{p}^*<1$.
%\end{enumerate}
\begin{enumerate} % gamma version
\item Fixation of cooperation,
	if $T_A>T_B$ and $c<\gamma_1$; or
	if $T_A>T_B$ and $\gamma_1<c<\gamma_2$ and $\tilde{p}_0>\tilde{p}^*$; or
	if $T_A<T_B$ and $c<\gamma_2$.
\item Fixation of defection,
	if $T_A>T_B$ and $\gamma_2<c$; or
	if $T_A>T_B$ and $\gamma_1<c<\gamma_2$ and $\tilde{p}_0<\tilde{p}^*$; or
	if $T_A<T_B$ and $\gamma_1<c$.
\item Coexistence of both phenotypes at $\tilde{p}^*$, 
	if $T_A<T_B$ and $\gamma_2<c<\gamma_1$.
\end{enumerate}

@@

@@

@@

@@

\begin{figure}[H]
  \centering
  \begin{subfigure}{8cm}
    \includegraphics[scale=0.5]{figure2a.pdf}
    \caption{$v=1$, $T_A=T_B=T$, $\alpha \neq 0$}
    \label{fig:results_a}
  \end{subfigure}
  \begin{subfigure}{8cm}
    \includegraphics[scale=0.5]{figure2b.pdf}
    \caption{$v=1$, $\alpha = 0$}
    \label{fig:results_b}
  \end{subfigure}
  \begin{subfigure}{8cm}
    \includegraphics[scale=0.5]{figure2c.pdf}
    \caption{$v=0$}
    \label{fig:results_c}
  \end{subfigure}
  \label{fig:results}
  \caption{
  \textbf{Numerical results for cultural evolution of cooperation.}
  Shown are dynamics of \textbf{(a-b)} $\tilde{p}$, the frequency of parents with cooperative phenotype $A$; \textbf{(c)} $p'$, the frequency of adults with cooperative phenotype $A$.
  The figure demonstrates fixation of cooperation (green), extinction of cooperation (blue)m and stable co-existence of cooperators and defectors (orange).
  %TODO ic (c), the blue and green got mixed. 
  % TODO have all panels in one row.
  }
\end{figure}

>>>>

We define the initial frequency as $\tilde{p}_0$ and the cost boundaries
\begin{equation} \label{eq:cost_boundaries}
\begin{aligned}
\gamma_1 = \frac{b \alpha T_A + (T_A - T_B)}{1-T_B}, \quad
\gamma_2 = \frac{b \alpha T_B + (1+b) (T_A - T_B)}{1-T_B}.
\end{aligned}
\end{equation}
Applying eqs.~\ref{eq:vert_hori_equilibrium}, \ref{eq:vert_hori_global_condition1}, and \ref{eq:vert_hori_global_condition2} we can summarize the possible outcomes (\autoref{fig:results2}):
%\begin{enumerate} % p* version
%\item Fixation of cooperation,
%	if $T_A>T_B$ and $0<\tilde{p}^*$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0>\tilde{p}^*$, or
%	if $T_A<T_B$ and $1<\tilde{p}^*$.
%\item Global fixation of defection,
%	if $T_A>T_B$ and $\tilde{p}^*>1$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0<\tilde{p}^*$, or
%	if $T_A<T_B$ and $\tilde{p}^*<0$
%\item Co-existence of both phenotypes at $\tilde{p}^*$, 
%	if $T_A<T_B$ and $0<\tilde{p}^*<1$.
%\end{enumerate}
\begin{enumerate} % gamma version
\item Fixation of cooperation,
	if $T_A>T_B$ and $c<\gamma_1$; or
	if $T_A>T_B$ and $\gamma_1<c<\gamma_2$ and $\tilde{p}_0>\tilde{p}^*$; or
	if $T_A<T_B$ and $c<\gamma_2$.
\item Fixation of defection,
	if $T_A>T_B$ and $\gamma_2<c$; or
	if $T_A>T_B$ and $\gamma_1<c<\gamma_2$ and $\tilde{p}_0<\tilde{p}^*$; or
	if $T_A<T_B$ and $\gamma_1<c$.
\item Coexistence of both phenotypes at $\tilde{p}^*$, 
	if $T_A<T_B$ and $\gamma_2<c<\gamma_1$.
\end{enumerate}

It is interesting to note that cooperation and defection can stably coexist if there his horizontal bias for defection and the cost of cooperation is large but not too large. The recurrence for this case is illustrated in \autoref{fig:coexistence}.

@@

%%% Figure: Result 2 Boundaries

\begin{figure}[htb]
  \centering
    \includegraphics[scale=1]{figure2a.pdf}
    \includegraphics[scale=1]{figure2b.pdf}
  \caption{\textbf{Vertical and horizontal transmission.} The figure illustrates combinations of horizontal bias, $T_A-T_B$, and cost of cooperation, $c$, that lead to either global fixation of cooperation (green), global fixation of defection (red), fixation of either cooperation or fixation depending on the initial frequencies (blue), or coexistence of cooperation and defection (yellow). The blue and orange curves show the cost boundaries $\gamma_1$ and $\gamma_2$ (\autoref{eq:cost_boundaries}).
  Here, benefit of cooperation is $b=1.3$, horizontal transmission of cooperation $T_A=0.4$, $c$ and $T_B$ vary on the y- and x-axes. \textbf{Left}: With $\alpha = 0.7 > 0$, coexistence is possible (yellow). \textbf{Right}: With $\alpha=0$, coexistence is not possible.}
    \label{fig:results2}
\end{figure}
%%%

%%% Figure: coexistence map
\begin{figure}[htb]
\centering
\includegraphics[scale=1]{coexistence.pdf}
\caption{\textbf{Stable coexistence between cooperation and defection.} The frequency of the cooperative phenotype $A$ among juveniles in the next generation $\hat{p}'$ is plotted in blue as a function of the frequency in the current generation $\hat{p}$ for $T_A<T_B$ and $\gamma_2<c<\gamma_1$. The line of $\hat{p}'=p$ is in orange. The curve and line intersect at the equilibrium $\hat{p}^*$. 
The blue curve is above the orange line$\hat{p}<\hat{p}^*$, so that the frequency increases towards $\hat{p}^*$. The blue curve is below the orange line when $\hat{p}>\hat{p}^*$, so that the frequency decreases towards $\hat{p}^*$.
}
\label{fig:coexistence}
\end{figure}

@@

\autoref{fig:results2}b demonstrates this these conditions.

@@

\begin{figure}[htb]
  \centering
    \includegraphics[scale=0.75]{figure3a.pdf}
%    \caption{$v=1$, $T_A=T_B=T$, $\alpha \neq 0$}
    \label{fig:results_a}
    \includegraphics[scale=0.75]{figure3b.pdf}
%    \caption{$v=1$, $\alpha = 0$}
    \label{fig:results_b}
    \includegraphics[scale=0.75]{figure3c.pdf}
%    \caption{$v=0$}
  \label{fig:results}
  \caption{
  \textbf{Numerical results for cultural evolution of cooperation.}
  Shown are dynamics of \textbf{(a-b)} $\tilde{p}$, the frequency of parents with cooperative phenotype $A$; \textbf{(c)} $p'$, the frequency of adults with cooperative phenotype $A$.
  The figure demonstrates fixation of cooperation (green), extinction of cooperation (blue), and stable coexistence of cooperators and defectors (orange).
  Parameters: (a) $v=1$, $T_A=T_B=T$, $\alpha \neq 0$; (b) $v=1$, $\alpha = 0$; (c) $v=0$.
  % TODO add all parameters.
  % TODO remove legends, increase fonts, explain in caption instead of legend.
  }
\end{figure}

@@
